Wait-free Queues with Polygarithmic Step Complexity

### 2022-08-03 article/psuedocode.tex
review findResponse()

%\begin{lemma}[Search Ranges]\label{search}
%  Preconditions of all invocation of \nf{BSearch} are satisfied.
%\end{lemma}
%\begin{proof}
%  
%Line 83: \nf{Get(i)} is called if the result of a dequeue is not null. The search is among all blocks in the root.
%
%Line 88: This search tries to find the ith enqueue, knowing that it is in the left child. Search is done over the left subblocks. The start and end of the range are followed by definition. Line 92 is the same.
%
%Line 101: Here, the goal is to find the superblock. We know the distance between answer and the \nf{super[i]} is at most $p$, since at most $p$ processes could die.
%
%\end{proof}

@@

\begin{lemma}\label{sizeCorrectness}
  \nf{root.blocks[b].\size}is the \size of the queue if the operations in the prefix for the $b$th block in the root are applied with the order of $L$.  
\end{lemma}
\begin{proof}
If the \size of a queue is greater than 0 then a \nf{Dequeue()} would decrease the \size of the queue, otherwise the \size of the queue remains 0. By definition  of $L$ in definition \ref{ordering} enqueue operations come before dequeue operations in a block. If the queue won't become empty after $b$th block we can compute the size of the queue after $b$th block using with this equality \nf{root.blocks[b].size= root.blocks[b-1].size+ root.blocks[b].sum\sub{enq}- root.blocks[b].sum\sub{deq}} (Line \ref{computeLength}). Otherwise \nf{root.blocks[b-1].size+ root.blocks[b].sum\sub{enq}- root.blocks[b].sum\sub{deq}} would become negative. Since there are more dequeue operations than the size of the queue. See table \ref{qhistory}.
\end{proof}

\begin{lemma} \label{numberOfNND}
If the operations are applied with the order of $L$, the number of non-null dequeues in the prefix for a block \nf{b} is \nf{b.sum\sub{enq}-b.\size}
\end{lemma}
\begin{proof}
If there are $enqs$ enqueue operations in the prefix for \nf{b}, then the size of the queue after  prefix for \nf{b} is \nf{enqs - non-null deqs} by Observation 35. The correctness of the feild \size is showed in Lemma \ref{sizeCorrectness}.
\end{proof}

\begin{lemma}\label{nullReturn}
$D_{root,b,i}$ returns \nf{null} iff \nf{root.blocks[b-1].\size+ root.blocks[b].num\sub{enq}- i <0}.
\end{lemma}

\begin{lemma}[Computing Response] \label{computeHead}
  Let $S$ be the state of an empty queue if the operations in $L$ are applied on it. \nf{FindResponse(b,i)} returns the element that is the the response to $D_{root,b,i}$ in $S$. If the queue is empty in $S$ then it returns \nf{null}.
\end{lemma}
\begin{proof}
First note that by Definition \ref{ordering} the linearization ordering  of operations will not change as new operations come so instead of talking about the linearization of operations before the $E_{root, b, i}$ we talk about what if the whole operation in the linearization are applied on a queue.

>>>>

%\begin{lemma}[Search Ranges]\label{search}
%  Preconditions of all invocation of \nf{BSearch} are satisfied.
%\end{lemma}
%\begin{proof}
%  
%Line 83: \nf{Get(i)} is called if the result of a dequeue is not null. The search is among all blocks in the root.
%
%Line 88: This search tries to find the ith enqueue, knowing that it is in the left child. Search is done over the left subblocks. The start and end of the range are followed by definition. Line 92 is the same.
%
%Line 101: Here, the goal is to find the superblock. We know the distance between answer and the \nf{super[i]} is at most $p$, since at most $p$ processes could die.
%
%\end{proof}
\begin{definition}
   Assume the operations in $L$ are applied on an empty queue. If element of \nf{enqueue e} is the response to  \nf{dequeue d} then we say \nf{R(d)=e}. If \nf{d} 's response id \nf{null} (queue is empty) then \nf{R(d)=null}.
\end{definition}

@@

\begin{lemma}\label{sizeCorrectness}
  \nf{root.blocks[b].\size}is the \size of the queue if the operations in the prefix for the $b$th block in the root are applied with the order of $L$.  
\end{lemma}
\begin{proof}
 need to say? :: If the \size of a queue is greater than 0 then a \nf{Dequeue()} would decrease the \size of the queue, otherwise the \size of the queue remains 0. By definition \ref{ordering} enqueue operations come before dequeue operations in a block in $L$.

We prove the claim by induction on \nf{b}. Base case \nf{b=0} is trivial since the queue is initialy empty and \nf{root.blocks[0].size=0}. For \nf{b=i} we are going to use the hypothesis for \nf{b=i-1}. If there are more than \nf{root.blocks[i-1].size+ root.blocks[i].sum\sub{enq}} dequeue operations in \nf{root.blocks[i]} then the queue would become empty after \nf{root.blocks[i]}. Otherwise we can compute the size of the queue after $b$th block using with this equality \nf{root.blocks[b].size= root.blocks[b-1].size+ root.blocks[b].sum\sub{enq}- root.blocks[b].sum\sub{deq}} (Line \ref{computeLength}). See Table \ref{qhistory} for an example of running some blocks of operations on an empty queue.
\end{proof}

\begin{lemma}[Duality of \#non-null dequeues and \nf{block.size}] \label{numberOfNND}
If the operations are applied with the order of $L$, the number of non-null dequeues in the prefix for a block \nf{b} is \nf{b.sum\sub{enq}-b.\size}
\end{lemma}
\begin{proof}
There are \nf{b.sum\sub{enq}} enqueue operations in the prefix for \nf{b}, then the size of the queue after the prefix for \nf{b} is \nf{\#enqs - \#non-null dequeues} in the prefix for \nf{b}, by Observation 35. So \nf{\#non-null dequeues} is \nf{b.sum\sub{enq}-b.\size}. The correctness of the \nf{block.\size}field is shown in Lemma \ref{sizeCorrectness}.
\end{proof}

\begin{lemma}\label{nullReturn}
\nf{R(D\sub{root,b,i})} is \nf{null} iff \nf{root.blocks[b-1].\size+ root.blocks[b].num\sub{enq}- i <0}.
\end{lemma}

\begin{lemma}[Computing Response] \label{computeHead}
\nf{FindResponse(b,i)} returns \nf{R(D\sub{root,b,i}).element}.
\end{lemma}
\begin{proof}
First note that by Definition \ref{ordering} the linearization ordering  of operations will not change as new operations come so instead of talking about the linearization of operations before the $E_{root, b, i}$ we talk about what if the whole operation in the linearization are applied on a queue.
